Viewing Computer Science through Citation Analysis

### 2020-03-30 Scientometrics/rebuttal/rebuttal_letter.tex

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}
%SetFonts

\title{SCIM-D-19-01397: Response to Initial Critique}
\author{George Chacko}
%\date{}							% Activate to display a given date or no date

\begin{document}
\maketitle
%\section{}
%\subsection{}

\vspace{30 mm}
\noindent Dear Dr. Gl{\"a}nzel:\\

Thank you very much for your message of March 6, 2020 in which it was communicated that the manuscript would benefit from major revisions.  We agree with this assessment and thank the reviewers for their constructive critique. Accordingly, I am (also on behalf of my co-authors) resubmitting (i) a revised manuscript with markup (ii) a clean copy with only the revisions. Lastly, I have addressed the comments of the reviewers in the remainder of this document.

I hope you will find that our revisions improve the manuscript. We would have responded sooner but the distractions of COVID-19 manifested themselves in a number of ways that impacted normal function. I also hope that you and your colleagues are safe and healthy.\\

\noindent Sincerely

\vspace{8 mm}

\noindent George Chacko, PhD

\newpage

\section*{Reviewer 1} The authors have extended the works of Salton and Bergmark on the computer science research. This study has potential. However, there are several concerns which reviewer wants to mention as follows:

\begin{enumerate}

\item Section 2 is really overwhelming with numerical values. Sometimes, its difficult to follow up the discussion. It would be easy for the reader if authors can make unique way of explaining the cleaning and merging they have done. May be in the form of a table to mention the numerical values and a flow chart to explain the flow of the process till the agglomerative clustering.

\emph{The reviewer makes a good point. We rewritten Section 2 (Materials and Methods) to include a paragraph that emphasizes flow while not eliminating any of the details we think are necessary. We think it's more readable now.}

\item Section-wise brief discussion is missing at the end of Introduction.

\emph{We have revised the Introduction to better track the flow of information through the document}

\item Quality of some images is really poor. Please improve them to make it much more visible.

\emph {We are a little puzzled by this comment, since the pdfs generated on our machines show acceptable image quality. However, we have regenerated all the images in pdf (and TIFF @600 dpi). In some cases, modified axis labels to make them more readable. Thank you for flagging this defect. In the event that the manuscript is accepted for publication, we will certainly work closely with journal staff to generate even higher resolution figures.}

\item Why is "Results" not a separate section?

\emph{This has been corrected, thank you. The error was on account of LaTeX formatting.}

\item Overall, the paper is interesting covering the field of computer science, which actually needs a separate attention, considering a lot of inter-disciplinary works fading away the actual computer science contribution. However, the presentation of this paper could be simplified for the broader audience.

\emph{Thank you for this comment, which is also raised by Reviewer 2. We have broadened the outreach in this revision.}
\end{enumerate}
\clearpage

\section*{Reviewer 2}

This manuscript overview the fields of computer science by the means of clustering for networks of direct citation or co-citation. Its originality is using DBLP bibliography, and I understand that the authors worked hard to process big data. As a result, I judge that this manuscript should be accepted. However, there are some points to be revised. (C01 - C05 are mandatory to be revised.)

\emph{Thank you for these comments. We would add respectfully that in terms of originality, our implementation of co-citation based clustering builds on rather than simply reproduces Small and Sweeney's approach.}

\begin{itemize}    
\item C01. The header ``Results" has a lack of numbering, and ``3 Discussion" will be ``4 Discussion".

\emph{Thank you for pointing this out. Our apologies, this has been fixed}

\item    C02. Some second-level headers in the sections ``2 Materials and Methods" and "Results" don't have period, so add period like "Overview.".

\emph{ Again, our thanks. We have corrected our previously sloppy editing for which we apologize}

\item C03. Update: "(Fig. 3, Cluster 17), ." $->$ "(Fig. 3, Cluster 17)."

\emph{A correction has been made}

\item C04. For the section ``1 Introduction," what is the purpose of this research, or what is this research aiming at? Ultimately, it is estimated that the part "we extend the work of Salton and Bergmark" will be the purpose. 
However, readers cannot understand the originality or significance of this research, because it seems to simply apply an existing methodology to current data. Thus, you should describe what you want to know from your 
results using the work of Salton and Bergmark. First, you should make the purpose specific in the section "1 Introduction". Second, answer to your new purpose in the section ``3 Discussion" or others (i.g. "Conclusion" if you add).
I recommend that you write an outspoken expression using "purpose" or "aim at" about your purpose for appealing your research's value to roughly-looking readers.

\emph{We interpret this comment from the reviewer as noting a lack of communication of purpose in the manuscript. We agree that in trying to reach out to multiple readerships with sometimes thinly overlapping interests, we have 
presented a broad sense of purpose. Therefore, we have tried to clarify purpose in the revision.}

\emph{Regarding `existing methodology to current data.', we will also offer that while we use Graclus, an existing methodology, for clustering by direct citation, it has not been used exhaustively in citation analysis as far as we can tell, and this is consistent with text in \v{S}ubelj, van Eck, and Waltman (2016). Further, we have modified co-citation based clustering as described by Small and Sweeney (1985). Lastly, that clustering methods tend to display sensitivity to input data such that every case has to be carefully evaluated. However, it is our responsibility to communicate this to the reader, so the reviewer's point is very well taken}

\item C05. How do the authors identify the field labels? The existing field labels would not be able to used, because your new borders by clustering and those of the exisiting field are different.

\emph{We are sorry that this was not clear. Yes, the new borders are different. Therefore, in each case we we only considered `Scopus ASJC minor subject area categories that accounted for at least 15\% of the publications in each cluster'. Thus, a cluster could map to multiple subject areas if at least 15\% of its content matched to each of them. We had to try setting different thresholds, indeed we show the heatmap for co-citation is shown at two different thresholds.}

\item C06. [Optional] In the section ``3 Discussion,"  you had better to write a limitation by excluding small fields, which may include innovative seeds. In order to use this research for the purpose of promoting academic-industry collaboration, you can also describe it as your future work. I like Fig.3 (c), because I'm excited to low score areas that may be new opportunities. Meanwhile, currently, agricultural IT is emerging and fishery starts to use IT. You may be able to provide information of new opportunities with your cutoff small fields by using figures like Fig.3 (c)

\emph{For reasons of scope and feasibility, we restricted ourselves to interpreting strong signals from a measurement at a single point in time. Further, `To focus on relatively high signal, we only considered Scopus ASJC minor subject area categories that accounted for at least 15\% of the publications in each cluster', but we have added a sentence to the Discussion about excluding small fields, which may (or may not) include innovative seeds}

\item C07. [Optional] As a significance of your research, the authors can suggest the academic about a question if some current systems that a paper connects to one field are outdated.

\emph{Even though we find the idea attractive, we are hesitant to verge into speculation. We do make the point that contextualized interpretation of these results is important and will address the suggestion to a greater extent in the revision}

\item I'm looking forward to reviewing the updated manuscript again.

\emph{Thank you!}

\end{itemize}

\end{document}

### 2020-03-30 Scientometrics/rebuttal/comp_revised.tex
updated source files

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestorew
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{comment}
\usepackage{hyperref}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

@@

\institute{Sitaram Devarakonda,  \at
              Netelabs, NET ESolutions Corporation, McLean, VA \\
              %Tel.: +123-45-678910\\
              %Fax: +123-45-678910\\
              \email{sitaram@nete.com}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
              Dmitriy Korobskiy \at
              Netelabs, NET ESolutions Corporation, McLean, VA  \\
              \email{dk@nete.com}
           \and
           Tandy Warnow \at
              Dept of Computer Science, University of Illinois Urbana-Champaign, Champaign IL \\
              \email{warnow@illinois.edu}
           \and
           George Chacko \at
              Netelabs, NET ESolutions Corporation, McLean, VA \\
              \email{netelabs@nwete.com}
}

@@

\begin{abstract}
Computer science has experienced dramatic growth and diversification over the last twenty years. Towards a current understanding of the structure of this discipline, we analyze 
a cohort of the computer science literature using the DBLP database. For insight on the features of this cohort and the relationship within its components, we constructed article 
level clusters based on either direct citations or co-citations, and reconciled them to major and minor subject categories in the All Science Journal Classification (ASJC). 
We described complementary insights from clustering by direct citation and co-citation, and both point to the increase in computer science publications and their scope.
Our analysis shows cross-category clusters, some that interact with external fields, such as the biological sciences, while others remain inward looking.

@@

Computer science, and its applications, has experienced rapid growth and diversification over the last twenty years. As observed in a 2017 US National Academies Report, ``A wide range of jobs in virtually all sectors demand computing skills to an unprecedented extent. And every academic discipline finds itself incorporating computing into its research and educational mission"~\cite{nas_2017}. More recently, the collective influence of the Internet of Things (IoT), `big' data, accessible cloud computing, and advances in artificial intelligence have been postulated as a recent driver for growth and evolution~\cite{siebel2019_digital}. Given this powerful influence, some understanding of the present state and structure of computer science and its relationship to other fields could inform planning and policy making at multiple levels from national level funding all the way down to faculty hiring strategy.

@@

In this study, we extend the work of Salton and Bergmark through a combination of article and journal approaches to study trends in computer science, while also considering connections to other fields, especially biology since approximately 42\% of Scopus is classified under the top level subject areas of Life Sciences and Health Sciences. As a source of computer  literature, we used DBLP, a reference bibliography for computer science~\cite{dblp_ref}. The DBLP bibliography covers publications from computer science and includes publications from hybrid fields, where they are considered pertinent to computer science research. We describe below the high-level interactions of computer science within the Physical Sciences, and with the Social Sciences, Life Sciences, and Health Sciences.

@@

 \emph{Overview} For the purpose of this study, our working definition of the computer science literature was all publications in the DBLP bibliography that (i) had a digital object identifier (DOI), and (ii) could also be matched to article identifiers in the Scopus bibliography. The underlying assumption is that records in DBLP are greatly enriched for computer science even while recognizing that coverage may be incomplete and possibly biased. Cross-matching DBLP publications to records in the Scopus abstract and citation database of peer-reviewed literature enabled us to harvest the richer links in Scopus as well as extract links to publications from other disciplines. This cross-matching to Scopus also allowed the use of both journal- and article-based classifications when clustering documents. We used DBLP articles from journals and conference proceedings to construct article clusters by using either direct citations or co-citations as links. We reconciled these clusters with the All Science Journal Classification (ASJC) developed and maintained by Scopus through a combination of automated and manual procedures, producing a dataset of  2,685,356 publications.

\emph{Data} A stable release of the DBLP computer science bibliography~\cite{dblp_ref} consisting of 7,079,994 records was downloaded as dblp-2018-08-01.xml.gz. 
Slightly over 95\% of the publications within were published after 1996. 
Publications were parsed from the XML source file and loaded into a PostgreSQL database. As part of implementing a larger data platform for research evaluation~\cite{GithubERNIE2019}, we have previously parsed the Scopus dataset, presently at over 88 million publications, into a custom schema in a PostgreSQL database. 
The total number of publications in Scopus labeled with major subject area Computer Science (in turn a subset of the Physical Sciences subject area) is 5,835,160. Records in the DBLP dataset were matched to Scopus identifiers using digital object identifiers (DOIs). This procedure resulted in a dataset of 2,685,356 DBLP publications with Scopus identifiers where 1,278,322 (47.6\%) were labeled as article and 1,407,034 (52.4\%) as conference proceedings.  References cited by these publications were then extracted from Scopus (7,129,006 records), resulting in a total of 8,000,411 
publications and references. 
We represented these data using a graph where the 8,000,411 nodes represent
publications and references  and the 44,296,381 undirected edges represent citations within the dataset. 
We refer to these data as the  \emph{comp} dataset (Fig.~\ref{fig:ar_cp_annotation}, Table~\ref{tab:comp}).

@@

We first calculated the number of citations accumulated across Scopus for all 2,685,356 articles in \emph{comp}. After discarding those publications without any citations, we restricted further analysis to highly cited articles--those in the 90th percentile of citations or higher--resulting in a dataset of size 212,311. We then identified 4,318,305 publications in Scopus that cite these 212,311 articles from \emph{comp}. For each of the 4,318,305 citing publications in turn, all possible ${n \choose 2}$ reference pairs were generated from a publication's cited references, where $n$ is the number of references in a publication. The cited reference pairs this generated were then restricted to those in the set of 212,311 highly cited papers previously identified. A total of 46,463,117 unique co-cited pairs were thus obtained. 
The frequency of these co-cited pairs was then computed across the \emph{comp} dataset  and normalized using Salton's cosine formula~\cite{salton_citation_1979} to limit dominance by areas with high citation activity. These data were represented in a graph where each node was a publication and the weighted edge between the pair was the normalized co-citation frequency. \par

@@

\section*{Results}
\label{sec:results}

 In our high-level study of the structure of a twenty year cohort of the computer science literature, we chose to use both traditional journal-based and article-based approaches. Considering that traditional disciplines `may only partly reflect the actual organization of today's scientific research'~\cite{waltman_new_2012}, we constructed article clusters at high levels of aggregation using citations to examine the  computer science literature and also mapped these clusters to journal-based categories to take advantage of both article level and journal level approaches.

\emph{Data} The process of selection and matching resulted in a dataset of  publications (Materials and Methods). Of 4,291,130 DBLP publications with DOIs, only  2,685,356 had corresponding DOIs in Scopus. Of these roughly 2.68 million publications, approximately 2.07 million were assigned ASJC codes in Scopus corresponding to Computer Science (major subject area with 13 minor subject areas (Table~\ref{tab:comp}), with the balance of 610,000 publications non-exclusively shared between 26 different major subject areas ranging from Engineering (330,048) to Dentistry (Fig.~\ref{fig:ar_cp_annotation}). The set of 2.07 million publications classified under the major subject area Computer Science in Scopus spanned all 13 minor subject areas with Software at 30.3\%  being the largest component and Computer Science (miscellaneous) at 0.9\% the smallest. Publications in the \emph{comp} dataset labeled Theoretical Computer Science (409,082) are classified under the the major subject area of Mathematics rather than Computer Science (Table~\ref{tab:comp}).

@@

At the threshold of 15\%, only 17 of 20 co-citation clusters mapped to at least one minor subject areas. At the 15\% threshold, no co-citation cluster maps to more than three minor subject areas, in contrast to a maximum of eight minor subject areas for clustering by direct citation (Fig. 3, Cluster 17), . The threshold had to be reduced to 10\% for all 20 co-citation clusters to map to at least one minor subject area. We interpret these data as indicative of broader clusters created by weaker linkages that accumulate during the agglomerative clustering phase~\cite{small_clustering_1985}.

@@

\section{Discussion} 
Considering expansion and diversification of the field of computer science, we revisited its characterization by Salton and Bergmark in 1979~\cite{salton_citation_1979}. In comparison to \cite{salton_citation_1979}, we analyzed considerably more data, 2.68 million publications versus 391 by using two bibliographic databases, DBLP and Scopus, consisting of  approximately 7 million and 88 million publications respectively. By linking the two datasets to harvest citation and metadata, we were able to construct article level clusters in two different ways and reconcile these clusters with the Scopus ASJC classification, which provides a journal level perspective.

@@

The focus of this article was on high-level features, and a preference for simplicity and intuitiveness in the choice of methods. This preference may address questions  regarding generalizability and the specific choices we made, such as the (i) use of the DBLP dataset matched to Scopus, which may not capture all aspects of computer science and its interactions with other fields, (ii) the basis for clustering, and (iii) mapping the results of this clustering against a classification designed around journals rather than individual articles. We believe that an approach that combines journal-level  with article-level analyses is useful for studies of this kind.

We speculate that biology, often dominant in bibliometric studies, is restricted to two clusters on account of (i) the focus of DBLP,  (ii) our use of normalized co-citations, and (ii) the threshold set for detection. Future research should, of course, include complementary investigations at finer levels of granularity and sensitivity using article-level and topic approaches that others have developed~\cite{glanzel_using_2017,boyack_clustering_2011,boyack_investigating_2017,boyack_classification_2014,sjogarde_granularity_2019,traag_louvain_2019}. We also refer readers to a related study of the computer science literature~\cite{chakraborty_role_2018} that is focused on evolving interdisciplinarity in computer science using data from Microsoft Academic Research and a 24 category classification of Computer Science.

@@

%%%%%%%%%
\section*{Figures}
%%%%%%%%%

\begin{figure}[ht]
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=0.6]{ar_cp_ratio.pdf}
% figure caption is below the figure
\caption{Summary of DBLP data cross-matched with Scopus. 2,685,356 publications from DBLP were cross-matched with Scopus and then grouped by the 27 major subject areas in the ASJC (Scopus) classification. The largest number of publications are contributed by Computer Science; Engineering; Mathematics; and then by Social Sciences; Decision Sciences; Physics and Astronomy; Medicine; and Biochemistry, Genetics, and Molecular Biology. Publications were further annotated with respect to being either articles \emph{(ar)} or conference proceedings \emph{(cp)}. For this dataset, the major subject area of Computer Science with 1,194,623(cp) \& 879,396 contributed the most publications while Dentistry with 0 (cp) \& 1 (ar)) contributed the least.}
\label{fig:ar_cp_annotation}       % Give a unique label
\end{figure}

@@

\begin{figure}[ht]
\centering
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=0.45]{scopus_dblp_graclus3.pdf}
% figure caption is below the figure
\caption{Heat map for the clustering obtained by direct citation. The  y-axis (rows) correspond to topics, defined by Scopus characterizations, and the
x-axis (columns) represent the 20 different clusters.  
Each cluster is characterized by topics that label at least 15\% of the publications in the cluster.
 }
\label{fig:heatmap}       % Give a unique label
\end{figure}

@@

\end{document}
% end of file template.tex

>>>>

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestorew
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{listings}
%\usepackage{comment}
\usepackage{hyperref}
\usepackage{changes}
\definechangesauthor[name={George Chacko}, color=orange]{gc}
%\setremarkmarkup{(#2)}
%\usepackage[firstpage]{draftwatermark}

%\usepackage{draftwatermark}
%\SetWatermarkText{Confidential Draft}
%\SetWatermarkScale{0.5}
%\SetWatermarkColor[gray]{0.5}

%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}

% examples of changes syntax
%This is \added[id=per,remark={we need this}]{new} text.
% This is \added[id=per,remark={has to be in it}]{new} text.
% This is \deleted[id=per,remark=obsolete]{unnecessary}text.
% This is \replaced[id=per]{nice}{bad} text.
% This is \added[remark={we need this}]{new} text.
% This is \added[remark={has to be in it}]{new} text.
% This is \deleted[remark=obsolete]{unnecessary}text.
% This is \replaced{nice}{bad} text.
% to generate a clean copy edit the \usepackage{changes} above to \usepackage[final]{changes}

\begin{document}

@@

\institute{Sitaram Devarakonda,  \at
              Netelabs, NET ESolutions Corporation, McLean, VA \\
              %Tel.: +123-45-678910\\
              %Fax: +123-45-678910\\
              \email{sitaramssd@gmail.com}           %  \\
              \emph{Present address: Randstad USA, Atlanta, GA} 
           \and
              Dmitriy Korobskiy \at
              Netelabs, NET ESolutions Corporation, McLean, VA  \\
              \email{dk@nete.com}
           \and
           Tandy Warnow \at
              Dept of Computer Science, University of Illinois Urbana-Champaign, Champaign IL \\
              \email{warnow@illinois.edu}
           \and
           George Chacko \at
              Netelabs, NET ESolutions Corporation, McLean, VA \\
              \email{netelabs@nete.com}
}

@@

\begin{abstract}w
Computer science has experienced dramatic growth and diversification over the last twenty years. Towards a current understanding of the structure of this discipline, we analyze 
a \replaced{large sample}{cohort} of the computer science literature from the DBLP database. For insight on the features of this cohort and the relationship within its components, we have constructed article 
level clusters based on either direct citations or co-citations, and reconciled them \replaced{with}{to} major and minor subject categories in the All Science Journal Classification (ASJC). 
We describe complementary insights from clustering by direct citation and co-citation. \deleted[remark=repositioned]{, and both point to the increase in computer science publications and their scope}.
Our analysis \replaced{reveals}{shows} cross-category clusters, some that interact with external fields, such as the biological sciences, while others remain inward looking.\added{Overall, we document an increase in computer science publications and their scope.}

@@

Computer science, and its applications, has experienced rapid growth and diversification over the last twenty years. As observed in a 2017 US National Academies Report, ``A wide range of jobs in virtually all sectors demand computing skills to an unprecedented extent. And every academic discipline finds itself incorporating computing into its research and educational mission"~\cite{nas_2017}. More recently, the collective influence of the Internet of Things (IoT), `big' data, accessible cloud computing, and advances in artificial intelligence have been \replaced{presented as a driver for digital transformation}{postulated as a recent driver for growth and evolution}~\cite{siebel2019_digital}. Given this \replaced{rapid growth and expansion}{powerful influence}, an updated understanding of the present state and structure of computer science and its relationship to other fields \replaced{can only inform planning and policy making at multiple levels from national level funding all the way down to faculty hiring strategy.}{could inform planning and policy making at multiple levels from national level funding all the way down to faculty hiring strategy.}

@@

\replaced{The purpose of this study is to develop an improved understanding of the structure of the field of computer science relative to the landmark study of Salton and Bergmark forty years ago. We extend their work by takimg advantage of modern bibliographic resources and clustering technologies and using a combination of article and journal approaches to study trends in the computer science literature. We also consider connections to other fields, especially biology, since approximately 42\% of Scopus is classified under the top level subject areas of Life Sciences and Health Sciences. As a source of computer  literature, we use DBLP, a reference bibliography for computer science~\cite{dblp_ref}. The DBLP bibliography covers publications from computer science and includes publications from hybrid fields, where they are considered pertinent to computer science research.} {In this study, we extend the work of Salton and Bergmark through a combination of article and journal approaches to study trends in computer science, while also considering connections to other fields, especially biology since approximately 42\% of Scopus is classified under the top level subject areas of Life Sciences and Health Sciences. As a source of computer  literature, we used DBLP, a reference bibliography for computer science~\cite{dblp_ref}. The DBLP bibliography covers publications from computer science and includes publications from hybrid fields, where they are considered pertinent to computer science research. We describe below the high-level interactions of computer science within the Physical Sciences, and with the Social Sciences, Life Sciences, and Health Sciences.}

We XXX XXXX XXXXX

@@

\emph{Overview} \added[id=gc,{remark=Reviewer 1}]{The underlying assumption is that, notwithstanding incomplete coverage, records in DBLP are greatly enriched for computer science~\cite{dblp_ref}. Therefore, the contents of DBLP are an excellent sample of the computer science literature to analyze. The workflow described in the following paragraphs of Materials and Methods, consists at a high-level of (i) merging the DBLP and Scopus datasets using digital object identifiers(DOIs) (ii) extracting cited references for the articles common to DBLP and Scopus (iii) clustering by direct citation and by our modification of co-citation based clustering. The outcome was a dataset of 8,000,411 publications.}

\replaced{Our working definition of the computer science literature, for the purpose of this study, was all publications in the DBLP bibliography that (i) had a digital object identifier DOI, and (ii) could also be matched to article identifiers in the Scopus bibliography.}{For the purpose of this study, our working definition of the computer science literature was all publications in the DBLP bibliography that (i) had a digital object identifier (DOI), and (ii) could also be matched to article identifiers in the Scopus bibliography.} Cross-matching DBLP publications to records in the Scopus abstract and citation database of peer-reviewed literature enables us to harvest the richer links in Scopus, as well as \added{to} extract links to publications from other disciplines. Cross-matching to Scopus also allows the use of journal-\added{based} classifications when clustering documents at the article-level. We used DBLP articles from journals and conference proceedings to construct article clusters by using either direct citations or co-citations as links. We reconciled these clusters with the All Science Journal Classification (ASJC) developed and maintained by Scopus through a combination of automated and manual procedures, producing a dataset of  2,685,356 publications, which when combined with cited references extracted from Scopus grew to 8,000,411 publications.

\added{\emph{DBLP data}}A stable release of the DBLP computer science bibliography~\cite{dblp_ref} consisting of 7,079,994 records was downloaded as dblp-2018-08-01.xml.gz. Slightly over 95\% of the publications within were published after 1996. Publications were parsed from the XML source file and loaded into a PostgreSQL database.

\added{\emph{Scopus data}} As part of implementing a larger data platform for research evaluation~\cite{GithubERNIE2019}, we have previously parsed the Scopus dataset, presently at over 88 million publications, into a custom schema in a PostgreSQL database. 
The total number of publications in Scopus labeled with major subject area Computer Science (in turn a subset of the Physical Sciences subject area) is 5,835,160.

\added{Merging and graph construction} Records in the DBLP dataset were matched to Scopus identifiers using digital object identifiers (DOIs). This procedure resulted in a dataset of 2,685,356 DBLP publications with Scopus identifiers where 1,278,322 (47.6\%) were labeled as article and 1,407,034 (52.4\%) as conference proceedings.  References cited by these publications were then extracted from Scopus (7,129,006 records), resulting in a total of 8,000,411 publications and references.

\added{We represented these  8,000,411 records, referred to as the  \emph{comp} dataset (Fig.~\ref{fig:ar_cp_annotation}, Table~\ref{tab:comp}), as a graph where the 8,000,411 nodes represent publications and references  and the 44,296,381 undirected edges represent citations within the dataset.}

@@

\added{We began by identifying highly cited articles in \emph{comp} and selected those in the 90th percentile (212,311 articles). We then identified ~4.3 million publications in Scopus that cite these 212,311 articles. For each of the 4.3 million citing publications in turn, all possible ${n \choose 2}$ reference pairs were generated from a publication's cited references, where $n$ is the number of references in a publication. The cited reference pairs this generated were then restricted to those where both members of a pair were in the set of 212,311 highly cited papers previously identified. A total of 46,463,117 unique co-cited pairs were thus obtained. The frequency of these co-cited pairs was then computed across the \emph{comp} dataset  and normalized using Salton's cosine formula~\cite{salton_citation_1979} to limit dominance by areas with high citation activity. These data were represented in a graph where each node was a publication and the weighted edge between the pair was the normalized co-citation frequency. \par }

@@

\section{Results}
\label{sec:results}

 In our high-level study of the structure of a twenty year sample of the computer science literature, we chose to use both traditional journal-based and article-based approaches. Considering that traditional disciplines `may only partly reflect the actual organization of today's scientific research'~\cite{waltman_new_2012}, we constructed article clusters at high levels of aggregation using citations to examine the  computer science literature and also mapped these clusters to journal-based categories to take advantage of both article level and journal level approaches.

\deleted{\emph{Data}The process of selection and matching resulted in a dataset of  publications (Materials and Methods). Of 4,291,130 DBLP publications with DOIs, only  2,685,356 had corresponding DOIs in Scopus.} Of 2.68 million publications in \emph{comp}, approximately 2.07 million were assigned ASJC codes in Scopus corresponding to Computer Science (major subject area with 13 minor subject areas (Table~\ref{tab:comp}), with the balance of 610,000 publications non-exclusively shared between 26 different major subject areas ranging from Engineering (330,048) to Dentistry (Fig.~\ref{fig:ar_cp_annotation}). The set of 2.07 million publications classified under the major subject area Computer Science in Scopus spanned all 13 minor subject areas with Software at 30.3\%  being the largest component and Computer Science (miscellaneous) at 0.9\% the smallest. Publications in the \emph{comp} dataset labeled Theoretical Computer Science (409,082) are classified under the the major subject area of Mathematics rather than Computer Science (Table~\ref{tab:comp}).

@@

At the threshold of 15\%, only 17 of 20 co-citation clusters mapped to at least one minor subject areas. At the 15\% threshold, no co-citation cluster maps to more than three minor subject areas, in contrast to a maximum of eight minor subject areas for clustering by direct citation (Fig. 3, Cluster 17). The threshold had to be reduced to 10\% for all 20 co-citation clusters to map to at least one minor subject area. We interpret these results as indicative of broader clusters created by weaker linkages that accumulate during the agglomerative clustering phase~\cite{small_clustering_1985}.

@@

\section{Discussion} 
\added{Considering expansion and diversification of the field of computer science, we revisited its characterization by Salton and Bergmark in 1979~\cite{salton_citation_1979}. In comparison to \cite{salton_citation_1979}, we analyzed considerably more data, 2.68 million publications versus 391 by using two bibliographic databases, DBLP and Scopus, consisting of  approximately 7 million and 88 million publications respectively. By linking the two datasets, we were able to harvest citation data as well as other metadata that enabled theo construct article level clusters in two different ways and reconciliation with the Scopus ASJC classification, for a journal level perspective.}

@@

The focus of this article was on high-level features, and a preference for simplicity and intuitiveness in the choice of methods. This preference may address questions  regarding generalizability and the specific choices we made, such as the (i) use of the DBLP dataset matched to Scopus, which may not capture all aspects of computer science and its interactions with other fields, (ii) the basis for clustering, and (iii) mapping the results of this clustering against a classification designed around journals rather than individual articles. We believe that an approach that combines journal-level  with article-level analyses is useful for studies of this kind. \added{However, we acknowledge that our approach would not detect emerging fields with few publications that could be important signs of innovation.}

We speculate that biology, often dominant in bibliometric studies, is restricted to two clusters on account of (i) the focus of DBLP,  (ii) our use of normalized co-citations, and (iii) the threshold set for detection. Future research should, of course, include complementary investigations at finer levels of granularity and sensitivity using article-level and topic approaches that others have developed~\cite{glanzel_using_2017,boyack_clustering_2011,boyack_investigating_2017,boyack_classification_2014,sjogarde_granularity_2019,traag_louvain_2019}. We also refer readers to a related study of the computer science literature~\cite{chakraborty_role_2018} that is focused on evolving interdisciplinarity in computer science using data from Microsoft Academic Research and a classification of Computer Science into 24 categories.

@@

%%%%%%%%%
\section{Figures}
{Figures}
%%%%%%%%%

\begin{figure}[ht]
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
\includegraphics[scale=0.75]{ar_cp_ratio.pdf}
% generated by ar_cp_ratio.R
% figure caption is below the figure
\caption{Summary of DBLP data cross-matched with Scopus. 2,685,356 publications from DBLP were cross-matched with Scopus and then grouped by the 27 major subject areas in the ASJC (Scopus) classification. The largest number of publications are contributed by Computer Science; Engineering; Mathematics; and then by Social Sciences; Decision Sciences; Physics and Astronomy; Medicine; and Biochemistry, Genetics, and Molecular Biology. Publications were further annotated with respect to being either articles \emph{(ar)} or conference proceedings \emph{(cp)}. For this dataset, the major subject area of Computer Science with 1,194,623(cp) \& 879,396 contributed the most publications while Dentistry with 0 (cp) \& 1 (ar)) contributed the least.}
\label{fig:ar_cp_annotation}       % Give a unique label
\end{figure}

@@

\begin{figure}[ht]
\centering
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
\includegraphics[scale=0.6]{scopus_dblp_graclus3.pdf}
% generated by scopus_dblp_graclus2.R
% figure caption is below the figure
\caption{Heat map for the clustering obtained by direct citation. The  y-axis (rows) correspond to topics, defined by Scopus characterizations, and the
x-axis (columns) represent the 20 different clusters.  
Each cluster is characterized by topics that label at least 15\% of the publications in the cluster.
 }
\label{fig:heatmap}       % Give a unique label
\end{figure}

@@

% Non-BibTeX users please use
\listofchanges
\end{document}
% end of file template.tex
